---
title: "errors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{internal_errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Error implementation

There are two types of errors:

- External (i.e., normal) errors: They appear because the **user** made a mistake.
- Internal errors: They appear because the **developer** made a mistake.

### Implementation of errors

- Function `cli::cli_abort()` is used.
- Specify argument `class` in `cli::cli_abort()` to directly refer to this error in testthat
- Use `rlang::try_fetch()` to try if a call works and provide a specific error message if it fails. Set `parent = NA` if the reported function ("Error in function()") should be the function in which the `try_fetch()` call is located. If the expression, i.e., `some_function()` in the example below throws an error message that you want to keep, set `parent = cnd`.

```{r}
#| eval: false

rlang::try_fetch(
    some_call() # use {} for multi-line
    error = function(cnd) {
      cli::cli_abort(
        "message",
        class = "errorclass",
        parent = NA
      )
    }
  )
```

- When throwing errors from checkmate functions, wrap them in `try_fetch()` as shown below with error message `cnd$message` and `parent = NA`

```{r}
#| eval: false

rlang::try_fetch(
    checkmate::assert_something(),
    error = function(cnd) {
      cli::cli_abort(
        c(
          "x" = "Invalid input for function `xx()`",
          "i" = cnd$message
        ),
        class = "errorclass",
        parent = NA
      )
    }
  )
```

- If you want to include multi-line error messages, use a named character vector with the names as the symbols you want to show (No name = no prefix, single space = indented item, * = bullet, > = arrow, v = green "tick" symbol, x = red cross, ! = exclamation mark, i = info). You can use `cnd$message` inside these named character vectors as well.



####- `call`: All functions unrelated to S7 classes provide `rlang::caller_env()`. For ####S7-related functions, it is more complicated: ????

- For internal errors, use `.internal = TRUE` when calling `cli::cli_abort()` 
- Internal error messages: xxxx


### Implementation of internal errors

## Internal errors

| ID | Level | Error |
|------------------------|------------------------|------------------------|
| Script group 1 | **Metadata-related scripts** |  |
| Script 1.1 | metadata.R |  |
| Script 1.2 | metadata_constructor.R |  |
| Function group 1.2.1 | YAML evaluation functions |  |
| Function 1.2.1.1 | `yaml.read()` |  |
| Function 1.2.1.2 | `yaml.str.input()` |  |
| Function 1.2.1.3 | `yaml.forgive.component.name()` |  |
| Function 1.2.1.4 | `yaml.str.component()` |  |
| Function 1.2.1.5 | `yaml.str.options()` |  |
| Function 1.2.1.6 | `yaml.str.var.list()` |  |
| Function 1.2.1.7 | `yaml.str.var.groups()` |  |
| Function 1.2.1.8 | `yaml.str.na.codes()` |  |
| Function 1.2.1.9 | `yaml.str.contras()` |  |
| Function 1.2.1.10 | `yaml.add.name()` |  |
| Script 1.3 | metadata_properties.R |  |
| Script 1.4 | metadata_process_input.R |  |
| Function group 1.4.1 | Helpers |  |
| Function 1.4.1.1 | `craft.call.var.list()` |  |
| Function 1.4.1.2 | `craft.call.var.groups()` |  |
| Function 1.4.1.3 | `extract.symbols.from.ast()` |  |
| Error **IE14131** |  | Argument `x` is not of type constant (as checked by `rlang::is_syntactic_literal()`), symbol, call, or pairlist. |
| Function 1.4.1.4 | `check.rules.format()` |  |
| Function 1.4.1.5 | `check.new.format()` |  |
| Function group 1.4.2 | Processing functions |  |
| Function 1.4.2.1 | `process.cats()` |  |
| Function 1.4.2.2 | `process.dict()` |  |
| Function 1.4.2.3 | `process.dict.rules()` |  |
| Function 1.4.2.4 | `process.new()` |  |
| Script 1.5 | metadata_validator.R |  |
| Script group 2 | **Workflow functions** |  |
| Script 2.1 | state_of_data.R |  |
| Function group 2.1.1 | state_of_the_data() |  |
| Function 2.1.1.1 | state_of_the_data() |  |
| Error **IE21111** |  | After a workflow function, the state of the data is incorrect, as not all variables are character, even though they have to be. |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
|  |  |  |
